---
title: "Branch Length Simulations for the 3-Taxon Species Tree"
author: "Laura Kubatko, Alexander Leonard, and Julia Chifman"
date: "8/11/2022"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This document contains code to reproduce the simulations in the manuscript "Identifiability of speciation times under the multispecies coalescent" by Laura Kubatko, Alexander Leonard, and Julia Chifman for the 3-taxon species tree.<br><br>

We set the initial seed here, so that the results presented in the manuscript can be reproduced exactly. If you would like to repeat the simulation study with a different set of randomly generated data, you need only change the seed in the chunk below.

```{r}
set.seed(30927)
```

## Part I: Estimation of speciation times

The notation in these functions follows that of Zhu and Yang (2021), which differs from that used in the main text. Specifically, we use the parameters $\tau_0$, $\tau_1$, and $\theta$ to compute the true probabilities of each site pattern. We then randomly generate $n$ site patterns according to these probabilities. These randomly generated site patterns are then used to compute estimates for various quantities. In each of the settings below, we perform the above process $10,000$ times, generating $10,000$ estimates for each quantity. The functions to carry out ths simulations are below (set echo=TRUE to see them in the rendered document).


```{r, echo=FALSE}
num.trials <- 10000

GetTrueProbs <- function(tau0, tau1, theta) {
  a0 <- exp(-8 * tau0 / 3) / (3 + 4 * theta)
  a1 <- exp(-8 * tau1 / 3) / (3 + 4 * theta)
  b <- exp(-4 * tau1 / 3) / (3 + 2 * theta)
  
  p0 <- (1 + 18 * a0 + 54 * a0 * b + 9 * a1) / 16
  p1 <- 3 * (1 - 6 * a0 - 18 * a0 * b + 9 * a1) / 16
  p2 <- 3 * (1 + 6 * a0 - 18 * a0 * b - 3 * a1) / 16
  p3 <- p2
  p4 <- 6 * (1 - 6 * a0 + 18 * a0 * b - 3 * a1) / 16
  
  c(p0, p1, p2, p3, p4)
}

SimulateFrequencies <- function(tau0, tau1, theta) {
  true.probs <- GetTrueProbs(tau0, tau1, theta)
  rmultinom(1, num.trials, true.probs) / num.trials
}

EstimateTau0 <- function(theta, freqs) {
  f0 <- freqs[1]
  f1 <- freqs[2]
  f2 <- freqs[3]
  f3 <- freqs[4]
  
  tau0_est <- -3 * log((3 + 4 * theta) * (4 * f0 + 2 * f2 + 2 * f3 - 1) / 9) / 8
  tau0_est
}

EstimateTau1 <- function(theta, freqs) {
  f0 <- freqs[1]
  f1 <- freqs[2]
  f2 <- freqs[3]
  f3 <- freqs[4]
  
  tau1_est <- -3 * log((3 + 4 * theta) * (4 * f0 + 4 * f1 - 1) / 9) / 8
  tau1_est
}

GetCovarianceMatrix <- function(freqs) {
  mat <- matrix(nrow=5, ncol=5)
  for (i in 1:5) {
    for (j in 1:5) {
      if (i == j) {
        mat[i,j] <- freqs[i] * (1 - freqs[i]) / num.trials
      } else {
        mat[i,j] <- -1 * freqs[i] * freqs[j] / num.trials
      }
    }
  }
  mat
}

GetExp0Variance <- function(theta, freqs) {
  M <- GetCovarianceMatrix(freqs)
  k <- 4/3 + 16*theta/9
  v0 <- c(k, 0, k/2, k/2, 0)
  
  var0 <- t(v0) %*% M %*% v0
  var0
}

GetExp1Variance <- function(theta, freqs) {
  M <- GetCovarianceMatrix(freqs)
  k <- 4/3 + 16*theta/9
  v1 <- c(k, k, 0, 0, 0)
  
  var1 <- t(v1) %*% M %*% v1
  var1
}

CalcInterval <- function(exp.var.est, tau.est) {
  exp.est <- exp(-8 * tau.est / 3)
  interval <- c(exp.est + 1.96 * sqrt(exp.var.est), exp.est - 1.96 * sqrt(exp.var.est))
  interval <- (-3/8) * log(interval)
  interval
}


RunSimulation <- function(tau0, tau1, theta) {
  true.probs <- GetTrueProbs(tau0, tau1, theta)
  true.exp0.var <- GetExp0Variance(theta, true.probs)
  true.exp1.var <- GetExp1Variance(theta, true.probs)
  
  tau0.ests <- c()
  tau1.ests <- c()
  exp0.var.ests <- c()
  exp1.var.ests <- c()
  tau0.intervals <- matrix(nrow=0, ncol=2)
  tau1.intervals <- matrix(nrow=0, ncol=2)
  tau0.in.interval.counter <- 0
  tau1.in.interval.counter <- 0
  
  for (i in 1:10000) {
    freqs <- SimulateFrequencies(tau0, tau1, theta)
    
    tau0.est <- EstimateTau0(theta, freqs)
    tau1.est <- EstimateTau1(theta, freqs)
    exp0.var.est <- GetExp0Variance(theta, freqs)
    exp1.var.est <- GetExp1Variance(theta, freqs)
    tau0.interval <- CalcInterval(exp0.var.est, tau0.est)
    tau1.interval <- CalcInterval(exp1.var.est, tau1.est)
    
    tau0.ests <- append(tau0.ests, tau0.est)
    tau1.ests <- append(tau1.ests, tau1.est)
    exp0.var.ests <- append(exp0.var.ests, exp0.var.est)
    exp1.var.ests <- append(exp1.var.ests, exp1.var.est)
    tau0.intervals <- rbind(tau0.intervals, tau0.interval)
    tau1.intervals <- rbind(tau1.intervals, tau1.interval)
    
    if (tau0 > tau0.interval[1] && tau0 < tau0.interval[2]) {
      tau0.in.interval.counter <- tau0.in.interval.counter + 1
    }
    
    if (tau1 > tau1.interval[1] && tau1 < tau1.interval[2]) {
      tau1.in.interval.counter <- tau1.in.interval.counter + 1
    }
  }
  
  tau0.in.interval.freq <- tau0.in.interval.counter / 10000
  tau1.in.interval.freq <- tau1.in.interval.counter / 10000
  
  list(tau0.ests, tau1.ests, exp0.var.ests, exp1.var.ests, tau0.intervals, tau1.intervals, tau0.in.interval.freq, tau1.in.interval.freq)
}

#xlab=paste("", x.label)
MakeHistogram <- function(values, true.value=NULL, title, x.label) {
  hist(values, nclass=10, main=paste("", title), xlab="", cex.lab=1.7, cex.axis=1.4, col="lightskyblue")
  lines(c(mean(values), mean(values)), c(0, 2147483647), lwd=6, col="gray")
  if (!is.null(true.value)) {
    lines(c(true.value, true.value), c(0, 2147483647), lwd=3)
  }
}

MakeScatterplot <- function(exp.estimates, variances, true.variance, title) {
  plot(1:length(variances), variances, pch=20, xlab="Simulation Replicate", ylab="Variance", main=paste("", title), cex.lab=1.7, cex.axis=1)
  lines(c(0,2147483647), c(var(exp.estimates), var(exp.estimates)), lwd=3, col="gray")
  lines(c(0,2147483647), c(true.variance, true.variance), lwd=3, col="red")
}

#xlab=paste("", x.label)
MakeIntervalPlot <- function(intervals, true.value, title, x.label) {
  plot(c(min(intervals), max(intervals)), c(0, 101), type='n', main=paste("", title), xlab="", ylab="", cex.lab=2.0)
  for (i in 1:100) {
    interval <- intervals[i,]
    lines(c(interval[1], interval[2]), c(i, i))
  }
  lines(c(true.value, true.value), c(0, 101), col="maroon")
}

MakeSummaries <- function(tau0, tau1, theta, results) {
  true.probs <- GetTrueProbs(tau0, tau1, theta)
  true.exp0.var <- GetExp0Variance(theta, true.probs)
  true.exp1.var <- GetExp1Variance(theta, true.probs)
  
  tau0.ests <- results[[1]]
  tau1.ests <- results[[2]]
  exp0.var.ests <- results[[3]]
  exp1.var.ests <- results[[4]]
  tau0.intervals <- results[[5]]
  tau1.intervals <- results[[6]]
  tau0.in.interval.freq <- results[[7]]
  tau1.in.interval.freq <- results[[8]]
  
  par(mfrow=c(1,2),mai=c(1,1,0.25,0.25))
  
  MakeHistogram(tau0.ests, tau0, "Estimated values of tau0", "tau0")
  MakeHistogram(tau1.ests, tau1, "Estimated values of tau1", "tau1")
  
  MakeHistogram(exp0.var.ests, true.exp0.var, "Estimated variances of exp(-8/3 tau0)", "exp(-8/3 tau0)")
  MakeHistogram(exp1.var.ests, true.exp1.var, "Estimated variances of exp(-8/3 tau1)", "exp(-8/3 tau1)")
  
  MakeScatterplot(exp(-8 * tau0.ests / 3), exp0.var.ests, true.exp0.var, "Estimated variances of exp(-8/3 tau0)")
  MakeScatterplot(exp(-8 * tau1.ests / 3), exp1.var.ests, true.exp1.var, "Estimated variances of exp(-8/3 tau1)")
  
  MakeIntervalPlot(tau0.intervals, tau0, "Confidence intervals for tau0", "tau0")
  MakeIntervalPlot(tau1.intervals, tau1, "Confidence intervals for tau1", "tau1")
  
  print("Proportion of 95% CIs including tau0:")
  print(tau0.in.interval.freq)
  print("Proportion of 95% CIs including tau1:")
  print(tau1.in.interval.freq)
}
```



The parametrization of Zhu and Yang (2021) differs from ours. We provide speciation times in coalescent units, while Zhu and Yang use mutation units. To convert coalescent units to mutation units, multiply by $\theta$. We list the simulation conditions in both units below.

### Simulation Setting #1: $\tau_1 = 1.0,\tau_0 = 2.0, \theta = 0.001, n = 100,000$

In mutation units, the parameter settings are: $\tau_1 = 0.001, \tau_0 = 0.002, \theta = 0.001, n = 100,000$


```{r, fig.width=10, fig.height=4, echo=FALSE}
tau0 <- 0.002
tau1 <- 0.001
theta <- 0.001
num.trials <- 100000

results <- RunSimulation(tau0, tau1, theta)
MakeSummaries(tau0, tau1, theta, results)
```

<br><br>

### Simulation Setting #2: $\tau_1 = 1.0, \tau_0 = 2.0, \theta = 0.005, n = 100,000$

In mutation units, the parameter settings are: $\tau_1 = 0.001, \tau_0 = 0.01, \theta = 0.005, n = 100,000$


```{r, fig.width=10, fig.height=4, echo=FALSE}
tau0 <- 0.01
tau1 <- 0.005
theta <- 0.005
num.trials <- 100000

results <- RunSimulation(tau0, tau1, theta)
MakeSummaries(tau0, tau1, theta, results)
```

<br><br>


### Simulation Setting #3: $\tau_1 = 1.0, \tau_0 = 1.5, \theta = 0.001, n = 100,000$

In mutation units, the parameter settings are: $\tau_1 = 0.001, \tau_0 = 0.0015, \theta = 0.001, n = 100,000$


```{r, fig.width=10, fig.height=4, echo=FALSE}
tau0 <- 0.0015
tau1 <- 0.001
theta <- 0.001
num.trials <- 100000

results <- RunSimulation(tau0, tau1, theta)
MakeSummaries(tau0, tau1, theta, results)
```

<br><br>


### Simulation Setting #4: $\tau_1 = 1.0, \tau_0 = 1.5, \theta = 0.005, n = 100,000$

In mutation units, the parameter settings are: $\tau_1 = 0.005, \tau_0 = 0.0075, \theta = 0.005, n = 100,000$


```{r, fig.width=10, fig.height=4, echo=FALSE}
tau0 <- 0.0075
tau1 <- 0.005
theta <- 0.005
num.trials <- 100000

results <- RunSimulation(tau0, tau1, theta)
MakeSummaries(tau0, tau1, theta, results)
```




### Simulation Setting #5: $\tau_1 = 1.0,\tau_0 = 2.0, \theta = 0.001, n = 10,000$

In mutation units, the parameter settings are: $\tau_1 = 0.001, \tau_0 = 0.002, \theta = 0.001, n = 10,000$


```{r, fig.width=10, fig.height=4, echo=FALSE}
tau0 <- 0.002
tau1 <- 0.001
theta <- 0.001
num.trials <- 10000

results <- RunSimulation(tau0, tau1, theta)
MakeSummaries(tau0, tau1, theta, results)
```

<br><br>

### Simulation Setting #6: $\tau_1 = 1.0, \tau_0 = 2.0, \theta = 0.005, n = 10,000$

In mutation units, the parameter settings are: $\tau_1 = 0.001, \tau_0 = 0.01, \theta = 0.005, n = 10,000$


```{r, fig.width=10, fig.height=4, echo=FALSE}
tau0 <- 0.01
tau1 <- 0.005
theta <- 0.005
num.trials <- 10000

results <- RunSimulation(tau0, tau1, theta)
MakeSummaries(tau0, tau1, theta, results)
```

<br><br>


### Simulation Setting #7: $\tau_1 = 1.0, \tau_0 = 1.5, \theta = 0.001, n = 10,000$

In mutation units, the parameter settings are: $\tau_1 = 0.001, \tau_0 = 0.0015, \theta = 0.001, n = 10,000$


```{r, fig.width=10, fig.height=4, echo=FALSE}
tau0 <- 0.0015
tau1 <- 0.001
theta <- 0.001
num.trials <- 10000

results <- RunSimulation(tau0, tau1, theta)
MakeSummaries(tau0, tau1, theta, results)
```

<br><br>


### Simulation Setting #8: $\tau_1 = 1.0, \tau_0 = 1.5, \theta = 0.005, n = 10,000$

In mutation units, the parameter settings are: $\tau_1 = 0.005, \tau_0 = 0.0075, \theta = 0.005, n = 10,000$


```{r, fig.width=10, fig.height=4, echo=FALSE}
tau0 <- 0.0075
tau1 <- 0.005
theta <- 0.005
num.trials <- 10000

results <- RunSimulation(tau0, tau1, theta)
MakeSummaries(tau0, tau1, theta, results)
```



## Part II: Hypothesis testing for species delimitation

```{r, echo=FALSE}
HypothesisTest <- function(tau, theta) {
  true.probs <- GetTrueProbs(0.001, tau, theta)
  test.stats <- c()
  null.hypothesis.rejected.counter <- 0
  
  for (i in 1:10000) {
    freqs <- SimulateFrequencies(0.001, tau, theta)
    tau1.est <- EstimateTau1(theta, freqs)
    exp1.est <- exp(-8 * tau1.est / 3)
    exp1.var.est <- GetExp1Variance(theta, freqs)
    
    test.stat <- (exp1.est - 1) / sqrt(exp1.var.est)
    test.stats <- append(test.stats, test.stat)
    
    if (abs(test.stat) > 1.96) {
      null.hypothesis.rejected.counter <- null.hypothesis.rejected.counter + 1
    }
  }
  
  power <- null.hypothesis.rejected.counter / 10000
  list(test.stats, power)
}
```

We now perform hypothesis tests for $\tau_1 = 0$, which corresponds to a scenario in which speciation has not yet occurred. For this purpose, we may use the test statistic
\begin{equation}
Z = \frac{\hat{x}_1 - 1}{\sqrt{\hat{var}(\hat{x}_1)}},
\end{equation}
where $x_1 = e^{-8 \tau_1 / 3}$. We consider only setting #1 - the other settings are analogous. We set $\tau_1 = 1.0, \tau_0 = 2.0, \theta=0.001, n= 100,000$ (corresponding to $\tau_1 = 0.001, \tau_0 = 0.002, \theta = 0.001, n=100,000$ in mutation units) and perform $10,000$ hypothesis tests. We also explore the power of this hypothesis test in the plot beneath the histograms by considering various values of $\tau_1$ between 0 and 1.00. We have done this for two different values of $\theta$, 0.001 and 0.005, as can be seen in the legend. Note that the x-axis is in mutation units.

```{r, fig.width=10, fig.height=4, echo=FALSE}
par(mfrow=c(1,3),mai=c(1,1,0.25,0.25))

num.trials <- 100000

results1 <- HypothesisTest(0, 0.005)
test1.stats <- results1[[1]]
MakeHistogram(values=test1.stats, title="tau1=0", x.label="Test statistic")

results2 <- HypothesisTest(0.0001, 0.005)
test2.stats <- results2[[1]]
MakeHistogram(values=test2.stats, title="tau1=0.1", x.label="Test statistic")

results3 <- HypothesisTest(0.001, 0.005)
test3.stats <- results3[[1]]
MakeHistogram(values=test3.stats, title="tau1=1.0", x.label="Test statistic")

taus = seq(0,0.001,0.00005)
powers1 <- c()
powers2 <- c()

for (curr.tau in taus) {
  results1 <- HypothesisTest(curr.tau, 0.001)
  powers1 <- append(powers1, results1[[2]])
  
  results2 <- HypothesisTest(curr.tau, 0.005)
  powers2 <- append(powers2, results2[[2]])
}

par(mfrow=c(1,1),mai=c(1,1,0.25,0.25))
plot(taus, powers1, type='l', lwd=3, col="lightskyblue", xlab=expression(paste(tau[1])), ylab="Power", cex.lab=1.7, cex.axis=1.4)
lines(taus, powers2, col="lightskyblue", lwd=3, lty=2)
legend(0.0025, 0.6, legend=c(expression(paste(theta,"=0.001")), expression(paste(theta,"=0.005"))), lty=c(1,2), col=c("lightskyblue", "lightskyblue"), lwd=c(3,3), cex=1.1)
```